---
title: "P8105_hw6_dm3175"
author: "Devon Morgan"
date: "11/20/2018"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

# Homework 6 Solutions
Homework 6 is focused on the topics included in the Linear Models module for p8105. 

# Problem 1
The next problem explores data gathered by the Washington Post on homicides in 50 large US cities. 

### Load and Clean the Data
Data was read in from a CSV posted on GitHub.

```{r}
url = "https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv"
homicide_data = read_csv(url, na = "Unknown") %>% 
  janitor::clean_names() 
```

Data cleaning steps were subsequently followed: 

```{r}
clean_homicide_data = homicide_data %>% 
  mutate(city_state = str_c(city, ", ", state),
         solved = ifelse(disposition == "Closed by arrest", 1, 0),
         victim_race = ifelse(victim_race == "White", "white", "non-white"),
         victim_race = fct_relevel(victim_race, "white"),
         victim_age = as.numeric(victim_age)) %>% 
  filter(!(city_state == "Dallas, TX" | city_state == "Phoenix, AZ" | city_state == "Kansas City, MO" | city_state == "Tulsa, AL"))
       
```

Cleaning steps:

*  Created a `city_state` variable combining city and state. 
*  Created a binary variable `solved` indicating if crime was solved. "1" signifies solved ("Closed by arrest"), and "0" signifies unsolved "Closed without arrest" or "Open/No arrest". 
*  Modified `victim_race` to have categories "white" and "non-white", with "white" as the reference category.
*  Converted `victim_age` to numeric.
*  Filtered out Dallas, TX, Phoenix, AZ and Kansas City, MO (incomplete data for variables of interest), and Tulsa, AL (data entry error). 
